Relative Stable Maps in Genus One via Radial Alignments

### 2017-12-18 genus-one-relative-maps.tex
Finished smoothing of comb component, without log structures

The result is a fibred surface $W^\prime \to \Aaff^1_t$ with general fibre isomorphic to $C_0$ and central fibre isomorphic to $C$. The total space of $W^\prime$ is no longer smooth, but this is fine; the projection to $\Aaff^1_t$ is still flat. The central fibre is a linearly trivial Cartier divisor:
\begin{equation*} W^\prime_0 = C_0 + C_1 + \ldots + C_k = 0 \in \Pic W^\prime \end{equation*}
Fix $i \in \{1,\ldots,k\}$; then $r_i C_i$ is Cartier, although $C_i$ may not necessarily be. Furthermore, since
\begin{equation*} l C_0 = - \sum_{i=1}^k l C_i = - \sum_{i=1}^k m_i (r_i C_i) \end{equation*}
in $A_1(W^\prime)$, it follows that $lC_0$ is Cartier. Finally, a local computation shows that
\begin{equation*} r_i C_i \cdot C_0 = 1 \end{equation*}
for $i \in \{1,\ldots,k\}$. Now, let $x_1,\ldots,x_n$ denote the marked points of $C$. These are smooth points of the central fibre $W^\prime_0$, and hence can be extended to Cartier divisors $\tilde{x}_1,\ldots,\tilde{x}_n$ on $W^\prime$. Consider the line bundle:
\begin{equation*} \tilde{L} = \OO_{W^\prime}(l C_0 + \Sigma_{j=1}^n \alpha_j \tilde{x}_j) \end{equation*}
on $W^\prime$. We claim that this gives a smoothing of the line bundle $L=f^*\OO(1)$ on $C$, i.e. that $\tilde{L}|_{W^\prime_0} = L$. We show this by first restricting $\tilde{L}$ to each of the components $C_i$ of $W^\prime_0 \cong C$. For $i \in \{1,\ldots,k\}$, we have
\begin{align*} \tilde{L}|_{C_i} & = \OO_{C_i} \left( (l C_0 \cdot C_i) q_i + \sum_{x_j \in C_i} \alpha_j x_j \right) = \OO_{C_i} \left( (l/r_i) q_i + \sum_{x_j \in C_i} \alpha_j x_j \right)\\
& = \OO_{C_i} \left( m_i q_i + \sum_{x_j \in C_i} \alpha_j x_j \right) = L|_{C_i} \end{align*}
while for $i=0$ we have:
\begin{align*} \tilde{L}|_{C_0} & = \OO_{C_0} \left( - \sum_{i=1}^k (l C_i \cdot C_0) q_i + \sum_{x_j \in C_0} \alpha_j x_j \right) = \OO_{C_0} \left( - \sum_{i=1}^k m_i q_i + \sum_{x_j \in C_0} \alpha_j x_j \right)  = L|_{C_0} \end{align*}
Finally the fact that $\tilde{L}|_{W^\prime_0} = L$ follows from the fact that the dual intersection graph of $C$ has genus zero.

@@

Now, to extend the sections $s_1,\ldots,s_N$, we simply check that they are unobstructed. The space containing the obstructions is:
\begin{equation*} \HH^1(C,L) \end{equation*} \marginpar{Is this true even when $C$ is reducible}
By taking the normalisation exact sequence for $C$, tensoring with $L$ and passing to cohomology, we obtain:

>>>>

The result is a fibred surface $W^\prime \to \Aaff^1_t$ with general fibre equal to $C_0$ and central fibre $W^\prime_0 \cong C$. The total space of $W^\prime$ is no longer smooth (its singular points are [BLAH]), but this is not a problem since the projection to $\Aaff^1_t$ is still flat. The central fibre is a linearly trivial Cartier divisor:
\begin{equation*} W^\prime_0 = C_0 + C_1 + \ldots + C_k = 0 \in \Pic W^\prime \end{equation*}
For $i \in \{1,\ldots,k\}$ we have that $r_i C_i$ is Cartier, although the same is not necessarily true of $C_i$. Furthermore, since
\begin{equation*} l C_0 = - \sum_{i=1}^k l C_i = - \sum_{i=1}^k m_i (r_i C_i) \end{equation*}
in $A_1(W^\prime)$, it follows that $lC_0$ is Cartier. Finally, a local computation shows that
\begin{equation*} r_i C_i \cdot C_0 = 1 \end{equation*}
for $i \in \{1,\ldots,k\}$. Now, let $x_1,\ldots,x_n$ denote the marked points of $C$. These are smooth points of the central fibre $W^\prime_0$, and hence can be extended to Cartier divisors $\tilde{x}_1,\ldots,\tilde{x}_n$ on $W^\prime$. Consider the line bundle:
\begin{equation*} \tilde{L} = \OO_{W^\prime}(l C_0 + \Sigma_{j=1}^n \alpha_j \tilde{x}_j) \end{equation*}
on $W^\prime$. We claim that this gives a smoothing of the line bundle $L=f^*\OO(1)$ on $C$, i.e. that $\tilde{L}|_{W^\prime_0} = L$. We show this by first restricting $\tilde{L}$ to each of the components $C_i$ of $W^\prime_0 \cong C$. For $i \in \{1,\ldots,k\}$, we have
\begin{align*} \tilde{L}|_{C_i} & = \OO_{C_i} \left( (l C_0 \cdot C_i) q_i + \sum_{x_j \in C_i} \alpha_j x_j \right) = \OO_{C_i} \left( (l/r_i) q_i + \sum_{x_j \in C_i} \alpha_j x_j \right)\\
& = \OO_{C_i} \left( m_i q_i + \sum_{x_j \in C_i} \alpha_j x_j \right) = L|_{C_i} \end{align*}
while for $i=0$ we have:
\begin{align*} \tilde{L}|_{C_0} & = \OO_{C_0} \left( - \sum_{i=1}^k (l C_i \cdot C_0) q_i + \sum_{x_j \in C_0} \alpha_j x_j \right) = \OO_{C_0} \left( - \sum_{i=1}^k m_i q_i + \sum_{x_j \in C_0} \alpha_j x_j \right)  = L|_{C_0} \end{align*}
Finally the fact that $\tilde{L}|_{W^\prime_0} = L$ follows from the fact that the dual intersection graph of $C$ has genus zero.

@@

In order to extend the sections $s_1,\ldots,s_N$, we simply check that they are unobstructed. The space containing the obstructions to extending the sections is:
\begin{equation*} \HH^1(C,L) \end{equation*}
\marginpar{Is this true even when $C$ is reducible}By taking the normalisation exact sequence for $C$, tensoring with $L$ and passing to cohomology, we obtain an exact sequence:
\begin{align*} 0 \to & \HH^0(C,L) \to \bigoplus_{i=0}^k \HH^0(C_i,L) \xrightarrow{\theta} \bigoplus_{i=1}^k L_{q_i} \to \\
\to & \HH^1(C,L) \to \bigoplus_{i=0}^k \HH^1(C_i,L) \to 0 \end{align*}
Now, each of $C_1,\ldots,C_k$ is isomorphic to $\PP^1$ and $L|_{C_i}$ has non-negative degree. Hence the map $\theta$ is surjective. Thus the map
\begin{equation*} \HH^1(C,L) \to \bigoplus_{i=0}^k \HH^1(C_i,L) \end{equation*}
is an isomorphism. But $\HH^1(C_i,L)=0$ for $i \in \{1,\ldots,k\}$ since $C_i \cong \PP^1$ and $L|_{C_i}$ has non-negative degree, and by Serre duality
\begin{equation*} \HH^1(C_0,L) \cong \HH^0(C_0, L^\vee \otimes \omega_{C_0}) = \HH^0(C_0,L^\vee) = 0 \end{equation*}
where the penultimate equality holds because $\operatorname{g}(C_0)=1$ and the last equality holds because $L|_{C_0}$ has \emph{strictly} positive degree (here we are using the fact that $f|_{C_0}$ is non-constant).

To conclude, we have a family $\tilde{C} = W^\prime$ of nodal curves and a map from this family to $\PP^N$
\bcd
\tilde{C} \ar[r,"\tilde{f}"] \ar[d,"\pi"] & \PP^N \\
\Aaff^1_t & \,
\ecd
such that when we restrict to $0 \in \Aaff^1_t$ we get $C \to \PP^N$ and such that the general fibre is an element of the nice locus.
